---
title: "Project"
author: "Nina Cunha"
date: "2026-03-29"
output: pdf_document
---

```{r}
library(hoopR)
library(dplyr)
library(ggplot2)
library(caret)
library(skimr)
library(caTools)
library(car)
```

```{r}
mbb_2026 <- load_mbb_team_box(seasons = c(2026))
mbb_2025 <- load_mbb_team_box(seasons = c(2025))
```
```{r}
# Using directly relevant statistics to dataset
variables <- c("team_home_away", "blocks", "field_goals_made", "free_throw_pct", "steals", "three_point_field_goals_made", "total_turnovers", "defensive_rebounds",
                  "field_goals_attempted", "free_throws_made", "team_turnovers", "three_point_field_goals_attempted", "turnover_points", "team_winner", "fast_break_points",
                  "flagrant_fouls","total_rebounds", "turnovers","assists", "field_goal_pct", "fouls", "points_in_paint")
mbb_stat_2026 <- mbb_2026[, variables]
mbb_stat_2025 <- mbb_2025[, variables]
```
```{r}
dim(mbb_stat_2025)
```
```{r}
str(mbb_stat_2025)
```
```{r}
head(mbb_stat_2025)
```



```{r, warning=FALSE}
# convert non numeric variables to numbers
mbb_stat_2026$fast_break_points <- as.numeric(mbb_stat_2026$fast_break_points)
mbb_stat_2026$team_home_away <- ifelse(grepl("home", mbb_stat_2026$team_home_away), 1, 0)
colnames(mbb_stat_2026)[colnames(mbb_stat_2026) == "team_home_away"] <- "home_team"
mbb_stat_2026$team_winner <- ifelse(mbb_stat_2026$team_winner == TRUE, 1, 0)
mbb_stat_2026$points_in_paint <- as.numeric(mbb_stat_2026$points_in_paint)
mbb_stat_2026$turnover_points <- as.numeric(mbb_stat_2026$turnover_points)

# Same for 2025 season
mbb_stat_2025$fast_break_points <- as.numeric(mbb_stat_2025$fast_break_points)
mbb_stat_2025$team_home_away <- ifelse(grepl("home", mbb_stat_2025$team_home_away), 1, 0)
colnames(mbb_stat_2025)[colnames(mbb_stat_2025) == "team_home_away"] <- "home_team"
mbb_stat_2025$team_winner <- ifelse(mbb_stat_2025$team_winner == TRUE, 1, 0)
mbb_stat_2025$points_in_paint <- as.numeric(mbb_stat_2025$points_in_paint)
mbb_stat_2025$turnover_points <- as.numeric(mbb_stat_2025$turnover_points)
```
```{r}
summary(mbb_stat_2025)
```
```{r}
aggregate(mbb_stat_2025$team_winner, by=list(mbb_stat_2025$home_team), mean)
```
* **Teams playing at home won 67% of their games**


```{r}
# remove variables that may cause multicollinearity issues
my_cor <- cor(mbb_stat_2026)
findCorrelation(my_cor, cutoff=0.8, verbose=TRUE)

# same for 2025 season
my_cor_2 <- cor(mbb_stat_2025)
findCorrelation(my_cor_2, cutoff=0.8, verbose=TRUE)
```
```{r}
# Dropping columns 17 and 18 from dataset to remove multicollinearity
mbb_stat_2026 <- mbb_stat_2026[, -17]
mbb_stat_2026 <- mbb_stat_2026[, -18]

mbb_stat_2025 <- mbb_stat_2025[, -17]
mbb_stat_2025 <- mbb_stat_2025[, -18]
```


```{r}
# contigency table for home team vs winner
mytable <- xtabs(~home_team + team_winner, data=mbb_stat_2026)
mytable2 <- xtabs(~home_team + team_winner, data=mbb_stat_2025)
mytable
```
```{r}
(test_2026 <- chisq.test(mytable))
(test_2025 <- chisq.test(mytable2))
```
**It can be determined that being the home team and being the winner are dependant on eachother**


```{r}
# train model on all of the variables in data set 
model <- glm(team_winner ~ ., family = "binomial", data=mbb_stat_2026)
summary(model)

model2 <- glm(team_winner ~ ., family = "binomial", data=mbb_stat_2025)
summary(model2)
```
* The coefficients dont change over each year
* shown that there are is a dependence between the hometeam and the winner of the game
* what other stat tests should I analyze / how would suggest remove # of variables (cross validation?)
* should I remove multicollinear columns before conducting any analysis like EDA or just before logistic regression
* train test splits?